Light inextensible strings under tension in the Schwarzschild geometry

### 2018-10-03 string_microarticle.tex
various rewriting tweaks

\begin{abstract}
%%%
Here I analyze the behaviour of stationary, light inextensible strings
under tension in the vicinity of a nonrotating black hole.  Proper
length (given by the spatial component of the Schwarzshild metric) is
minimized by the Euler-Lagrange equations.  Light inextensible taut
strings may assume topologically distinct configurations, and here I
give some numerical solutions and theoretical observations.  
%%%%
\end{abstract}

\begin{keyword}
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\MSC 83C57\sep 35Q75
%% Keywords
\KWD Black Holes\sep Light inextensible string
\end{keyword}

@@

%% main text
\section{Introduction}
\label{intro}
Light inextensible strings are a stalwart feature of elementary
physics.  A light string is not subject to gravity, and being under
tension means that its configuration is an extremum of proper length.
Of course, no real string is perfectly weightless or inextensible.
With tension~$T$, gravitational acceleration $g$ and (proper) mass per
(proper) unit length~$\mu$, then a string will behave as though light
on lengthscales $L\lesssim\frac{T}{\mu g}$.  Using Newtonian
expressions for the strength of gravity at the event horizon of a
black hole gives $L\lesssim\frac{TGM}{\mu c^2}$.  For the string to
behave as though light on lengthscales comparable with the
Schwarzschild radius would require $T\gtrsim\mu c^2$.  Exceeding this
criterion would violate the dominant energy condition (and in this
context we note that transverse waves would propagate at
speed~$\sqrt{T/\mu}\sim c$).  Nevertheless, the idealization of a
light inextensible string is an interesting and informative limit: its
analysis interprets the concept ``straight line'' and suggests a
possible way of manipulating a black hole.  The Schwarzschild metric

@@

\noindent where units in which~$2GM=c=1$ are used, has been known for
over a century~\cite{schwarzschild1916} and makes a number of
counter-intuitive predictions.  Here I consider the spatial
component~$ds^2= \frac{dr^2}{1-1/r} + r^2d\phi^2$, confined to
equatorial points~($\theta=\pi/2$); we understand~$1<r<\infty$
throughout.  A string of constant~$r$ will, trivially, have proper
length~$2\pi r$.

@@

Light inextensible strings under tension are one interpretation of
``straight line'' in general relativity, the others being null
geodesics and world lines of massive objects.  Here it is shown that
such strings adopt non-trivial configurations including
self-intersecting paths.

In Figure~\ref{closest_approach_non_self_intersecting}, one's
classical intuition would suggest that the black hole feels a force to
the left, and should start accelerating.  However, it is not clear by
what mechanism such a force would be transmitted: the situation is
steady, so no gravitational waves are produced; the string is
inextensible, so stores no energy.  This work considered only the
Schwarzschild metric and generalizing it to the Kerr
metric~\cite{kerr1963} might be interesting.

@@

Calculations are performed using the R programming language; source
code for the figures is available at

\url{https://github.com/RobinHankin/string.git}

>>>>

\begin{abstract}
%%%
Here I analyze the behaviour of stationary, light inextensible strings
under tension in the vicinity of a nonrotating black hole.  Such
strings minimize their proper length, given by integrating the spatial
equatorial component of the Schwarzshild metric.  The string path is
calculated by numerical solution to the Euler-Lagrange equations.  If
the string is entirely outside the event horizon, its closest approach
is to the singularity is tangential; if the distance of closest
approach is less than about $1.076109317\cdot(2GM)$, the string
self-intersects.  Light taut strings interpret the concept ``straight
line'' and their behaviour suggests a possible way of manipulating a
black hole.
%%%%
\end{abstract}

\begin{keyword}
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\MSC 83C57\sep 35Q75
%% Keywords
\KWD Black holes\sep Light inextensible string
\end{keyword}

@@

%% main text
\section{Introduction}
\label{intro}
Light inextensible strings are a stalwart feature of elementary
physics.  A light string is not subject to gravity, and being under
tension means that its configuration is an extremum of proper length.
Of course, no real string is perfectly weightless or inextensible.
With tension~$T$, gravitational acceleration $g$ and (proper) mass per
(proper) unit length~$\mu$, then a string will behave as though light
on lengthscales $L\lesssim T/\left(\mu g\right)$.  Using Newtonian
expressions for the strength of gravity at the event horizon of a
black hole gives $L\lesssim TGM/\left(\mu c^2\right)$.  For the string
to behave as though light on lengthscales comparable with the
Schwarzschild radius would require $T\gtrsim\mu c^2$.  Exceeding this
criterion would violate the dominant energy condition (and in this
context we note that transverse waves would propagate at
speed~$\sqrt{T/\mu}\sim c$).  Nevertheless, the idealization of a
light inextensible string is an interesting and informative limit: its
analysis interprets the concept ``straight line'' and suggests a
possible way of manipulating a black hole.  The Schwarzschild metric

@@

\noindent where units in which~$2GM=c=1$ are used, has been known for
over a century~\cite{schwarzschild1916} and makes a number of
counter-intuitive predictions.  Here I consider the spatial
component~$ds^2= dr^2/\left(1-1/r\right) + r^2d\phi^2$, confined to
equatorial points~($\theta=\pi/2$); we understand~$1\leq r<\infty$
throughout.  A circular string at constant~$r$ will, trivially, have
proper length~$2\pi r$.

@@

Light inextensible strings under tension adopt non-trivial
configurations including self-intersecting paths.  The path of such
strings furnishes an interpretation of the concept ``straight line''
in general relativity, others being null geodesics and world lines of
massive objects.  Comparing a world line with that of nearby taut
strings thus furnishes another insight into path curvature,
complementing the analysis of Abramowicz~\cite{abramowicz1992}.

In Figure~\ref{closest_approach_non_self_intersecting}, one's
classical intuition would suggest that the black hole feels a force to
the left, and should start accelerating.  However, it is not clear by
what mechanism such a force would be transmitted: the situation is
steady, so no gravitational waves are produced, and the string is
inextensible so stores no energy.

@@

Calculations are performed using the R programming language; source
code for the figures is available at
\\
\\
\url{https://github.com/RobinHankin/string.git}
